---
title: "Adding Lat and Long to Case Data"
author: "Whitney Mgbara"
date: "4/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Libraries
```{r}

require(tidyverse)
require(readr)

require(ggplot2)

require(ggmap)
require(sf) 
require(rgdal)
require(raster)
```



US cities and county data
```{r}
# Counties
county_lat_lon <- read.csv("county_lat_lon.csv", header=TRUE, stringsAsFactors=F)
county_lat_lon <- dplyr::select(county_lat_lon, -c("ID", "County", "County.Seat.s."))
names(county_lat_lon)[names(county_lat_lon)=="Longitude"] <- "lon"
names(county_lat_lon)[names(county_lat_lon)=="Latitude"] <- "lat"

# Cities
uscities <- read.csv("uscities.csv")
names(uscities)[names(uscities)=="county_fips"] <- "FIPS"
names(uscities)[names(uscities)=="ST"] <- "State"

```


Join with case data and plot
```{r}
day25 <- read.csv("day25.csv", header=TRUE, stringsAsFactors=F)


```


```{r}

# Matching counties in day25 to lat/lon in county_lat_lon
geoday25_1 <- left_join(day25, county_lat_lon, by = c("FIPS", "State")) 



# Matching every city in uscities to a county in day25
geoday25_2 <- left_join(day25, uscities, by = c("FIPS", "State")) %>%
  dplyr::select("FIPS", "Name", "city", "State", "state_name", 
                "lat", "lng", "timezone", "zips")
names(uscities)[names(uscities)=="State"] <- "ST"



```



Match up day25 counties to airport enplanement data
```{r}

# Can call Rmarkdown file in same folder to create
# firstAirportData.csv
# source("airports_cities_counties.Rmd")

enplane <- read.csv("cy18-commercial-service-enplanements.csv")
enplane <- filter(enplane, CY.18.Enplanements != "NA")

test <- merge(enplane, uscities)
test$CY.18.Enplanements <- as.numeric(test$CY.18.Enplanements)
test$CY.17.Enplanements <- as.numeric(test$CY.17.Enplanements)
test$X..Change <- as.numeric(test$X..Change)

df1 <- 
  test %>% 
  group_by(FIPS) %>%
  summarize(CY.18.Enplanements =
              sum(CY.18.Enplanements),
            CY.17.Enplanements = sum(CY.17.Enplanements),
            X..Change = sum(X..Change)
            )

test1 <- left_join(geoday25_1, df1)


df2 <- 
  test %>% 
  group_by(FIPS, IATA) %>%
  summarize(CY.18.Enplanements =
              sum(CY.18.Enplanements),
            CY.17.Enplanements = sum(CY.17.Enplanements),
            X..Change = sum(X..Change)
            )

test2 <- left_join(geoday25_1, df2)
```

Data set with all airplane lat/lon and day25 county lat/lon 
```{r}
usairports <- read.csv("us-airports.csv")
names(usairports)[names(usairports)=="municipality"] <- "city"
usairports <- dplyr::select (usairports,-c(continent))

trial <- merge(usairports, enplane, by = c("IATA", "city"))

```


Save final result
```{r}
# head(airport_complete4)

write.csv(trial,"MyairportData.csv", row.names = FALSE)

# write.csv(geoday25_2,"Counties_Allcities.csv", row.names = FALSE)

write.csv(geoday25_1,"Mygeday25Data.csv", row.names = FALSE)
```


Plot the new dataset with cases
```{r}
qmplot(lon, lat, data = geoday25_1, colour = I('red'), size = I(0.5), darken = .3)
```



Plot all airports
```{r}
qmplot(longitude_deg, latitude_deg, data = trial, colour = I('red'), size = I(0.5), darken = .3)
```







Plotting the lat and long for counties based on the range of values for Day of First Case
```{r}

#Some sample data
x <- sort(runif(100))
dat <- data.frame(x = x,y = x^2 + 1)
# Some external vector for the color scale
col <- sort(rnorm(100))

qplot(x, y, data=dat, colour=col) + scale_colour_gradient(low="red", high="blue")

#Some sample data
x <- sort(runif(100))
dat <- data.frame(x = x,y = x^2 + 1)
# Some external vector for the color scale
col <- sort(geoday25_1$DayOfFirstCases)

qplot(lon, lat, data=geoday25_1, colour=col) + scale_colour_gradient(low="red", high="blue")

```




```{r}

partyUSA <- st_read("C:/Users/Whitney/Downloads/USA_adm/USA_adm1.shp")

# partyUSA <- st_read("C:/Users/Whitney/Downloads/cb_2018_us_csa_20m")


plot(partyUSA[7])
```



Averages By metropolitan statistical area
```{r}

metroAreas <- read.csv("metro_stat_area.csv", header=TRUE, stringsAsFactors=F)


test123 <- left_join(geoday25_1, metroAreas, by = "Name") %>%
  filter(State.x == c("NY", "CA")) %>%
  subset(Core_Based_Statistical_Area != "none")

test123$popland <- test123$Population/test123$LandArea


df <- 
  test123 %>% 
  group_by(Core_Based_Statistical_Area, State.x) %>%
  summarize(DayOfFirstCases = mean(DayOfFirstCases),
            ConfirmedCasesDay25 = sum(ConfirmedCasesDay25),
            CommutingByPublicTransportation =
              mean(CommutingByPublicTransportation), 
            GDP = mean(GDP),
            popland = mean(popland))

df


```



```{r}

testing123 <- data.frame(
  y = log(test123$ConfirmedCasesDay25), 
  x = log(test123$CommutingByPublicTransportation)) 

sp <- ggplot(testing123, aes(x=x, y=y)) +
  geom_point() +
  theme(legend.title = element_blank())
sp 

```



```{r}

plot(testing123$x)
```


