---
title: "Adding Lat and Long to Case Data"
author: "Whitney Mgbara"
date: "4/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Libraries
```{r}

require(tidyverse)
require(readr)

require(ggplot2)

require(ggmap)
require(sf) 
require(rgdal)
require(raster)

require(airportr)
# require(dplyr)
# library(tibble)
```


## County lat/lon

US cities and county data
```{r}
# Counties
county_lat_lon <- read.csv("county_lat_lon.csv", header=TRUE, stringsAsFactors=F)
county_lat_lon <- dplyr::select(county_lat_lon, -c("ID", "County", "County.Seat.s."))
names(county_lat_lon)[names(county_lat_lon)=="Longitude"] <- "lon"
names(county_lat_lon)[names(county_lat_lon)=="Latitude"] <- "lat"

```


Join with case data and plot
```{r}
day25 <- read.csv("day25.csv", header=TRUE, stringsAsFactors=F)


```


combine case data and lat/lon data by county
```{r}

# Matching counties in day25 to lat/lon in county_lat_lon
geoday25_1 <- left_join(day25, county_lat_lon, by = c("FIPS", "State")) 


```







Plot the new dataset with cases
```{r}
qmplot(lon, lat, data = geoday25_1, colour = I('red'), size = I(0.5), darken = .3)
```




Plotting the lat and long for counties based on the range of values for Day of First Case
```{r}

#Some sample data
x <- sort(runif(100))
dat <- data.frame(x = x,y = x^2 + 1)
# Some external vector for the color scale
col <- sort(rnorm(100))

qplot(x, y, data=dat, colour=col) + scale_colour_gradient(low="red", high="blue")

#Some sample data
x <- sort(runif(100))
dat <- data.frame(x = x,y = x^2 + 1)
# Some external vector for the color scale
col <- sort(geoday25_1$DayOfFirstCases)

qplot(lon, lat, data=geoday25_1, colour=col) + scale_colour_gradient(low="red", high="blue")

```




Averages By metropolitan statistical area
```{r}

metroAreas <- read.csv("metro_stat_area.csv", header=TRUE, stringsAsFactors=F)


test123 <- left_join(geoday25_1, metroAreas, by = "Name") %>%
  filter(State.x == c("NY", "CA")) %>%
  subset(Core_Based_Statistical_Area != "none")

test123$popland <- test123$Population/test123$LandArea


df <- 
  test123 %>% 
  group_by(Core_Based_Statistical_Area, State.x) %>%
  summarize(DayOfFirstCases = mean(DayOfFirstCases),
            ConfirmedCasesDay25 = sum(ConfirmedCasesDay25),
            CommutingByPublicTransportation =
              mean(CommutingByPublicTransportation), 
            GDP = mean(GDP),
            popland = mean(popland))

df


```


## PLotting county lat/lon with shapefiles
```{r}

partyUSA <- st_read("C:/Users/Whitney/Downloads/USA_adm/USA_adm1.shp")

# partyUSA <- st_read("C:/Users/Whitney/Downloads/cb_2018_us_csa_20m")


plot(partyUSA[7])
```





## Airplane lat/lon
us cities and counties
```{r}
# Cities
# getwd()
uscities <- read_csv("uscities.csv") 

uscities2 <- uscities %>%
  dplyr::select( c("city", "state_name", "county_fips", "county_name"))

names(uscities2)[names(uscities2)=="county_fips"] <- "FIPS"
names(uscities2)[names(uscities2)=="state_name"] <- "State"
names(uscities2)[names(uscities2)=="city"] <- "Location"
```


us commerical enplanement data
```{r}
enplane <- read.csv("cy18-commercial-service-enplanements.csv")
```


top us airports 
```{r}
air1 <- read.csv("topairports.csv")
air1$IATA <- as.character(air1$IATA)

# all us airports
air2 <- read.csv("us-airports.csv")
air3 <- data.frame(state_abrev = air2$iso_region, Location = air2$municipality, IATA = air2$IATA)
```



Get lat lon for top us airports 
```{r}

air1_lat_lon <- tibble()
# Needs airportr library
for (i in 1:nrow(air1)) {
 test <- airport_location(air1$IATA[i],input_type ="IATA")
 
 air1_lat_lon <- rbind(air1_lat_lon, test)

}


```


Combine airport data
```{r}
aye <- cbind(air1, air1_lat_lon) %>%
  left_join(air3, by = c("Location", "IATA")) %>%
  left_join(enplane, by = "IATA") %>%
  left_join(uscities2, by = c("Location", "State")) %>%
  dplyr::select(c("Location", "State",
                  "state_abrev", "county_name", "FIPS",
                  "Airport.Name", "IATA", "Rank", "Hub",
                  "Latitude", "Longitude",
                  "CY.18.Enplanements", "CY.17.Enplanements",
                  "X..Change"))

inday25 <- data.frame(FIPS = day25$FIPS,
                      # county_name = day25$Name, 
                      state_abrev = day25$State,
                      inday25 = 1)

aye2 <- left_join(aye, inday25, by = c("FIPS", "state_abrev"))
aye2$inday25 <- aye2$inday25 %>% replace_na(0)

```



```{r}

get_geo_distance = function(long1, lat1, long2, lat2, units = "miles") {
  loadNamespace("purrr")
  loadNamespace("geosphere")
  longlat1 = purrr::map2(long1, lat1, function(x,y) c(x,y))
  longlat2 = purrr::map2(long2, lat2, function(x,y) c(x,y))
  distance_list = purrr::map2(longlat1, longlat2, function(x,y) geosphere::distHaversine(x, y))
  distance_m = list_extract(distance_list, position = 1)
  if (units == "km") {
    distance = distance_m / 1000.0;
  }
  else if (units == "miles") {
    distance = distance_m / 1609.344
  }
  else {
    distance = distance_m
    # This will return in meter as same way as distHaversine function. 
  }
  distance
}
```



Plot top us airports airports
```{r}
qmplot(Longitude, Latitude, data = aye, colour = I('red'), size = I(3), darken = .3)
```




Data set with all airplane lat/lon and day25 county lat/lon 
```{r}

usairports2 <- air2
names(usairports2)[names(usairports2)=="municipality"] <- "city"

trial <- merge(usairports2, enplane, by = c("IATA", "city"))


qmplot(longitude_deg, latitude_deg, data = trial, colour = I('red'), size = I(0.5), darken = .3)
```



Save data
```{r}
# head(airport_complete4)

write.csv(trial,"MyairportData.csv", row.names = FALSE)

# write.csv(geoday25_2,"Counties_Allcities.csv", row.names = FALSE)

write.csv(geoday25_1,"Mygeday25Data.csv", row.names = FALSE)
```



