---
title: "COVID_day25_SL"
author: "Whitney Mgbara"
date: "4/15/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries
```{r libraries, include=FALSE}
library(here)
library(dplyr)
library(SuperLearner)
library(tmle3)
library(varimpact)
library(readxl)
library(readr)
library(data.table)
library(tidyverse)
library(sl3)
library(sjPlot)
library(ggplot2)
library(car)
library(msm)
library(MASS)

data(efc)
theme_set(theme_sjplot())

```

Import case data and covariates
```{r import, error=FALSE, message=FALSE, warning=FALSE}
# Case data
day25 <- read_excel("Data/day25.xls")
day25$popland <- day25$Population/(day25$LandArea)

# Covariate Data
analytic_data2020 <- read_csv("Data/analytic_data2020.csv")
colnames(analytic_data2020)[which(names(analytic_data2020) == "State Abbreviation")] <- "State"

eco_vars_of_int <- c("State", 
                     "Name",
                     "Poor or fair health raw value", 
                     "Adult smoking raw value",
                     "Food environment index raw value", 
                     "Physical inactivity raw value",
                     "Excessive drinking raw value", 
                     "Sexually transmitted infections raw value",
                     "Primary care physicians raw value", 
                     "Flu vaccinations raw value",
                     "High school graduation raw value", 
                     "Unemployment raw value", 
                     "Air pollution - particulate matter raw value", 
                     "Drinking water violations raw value",
                     "Diabetes prevalence raw value", 
                     "HIV prevalence raw value", 
                     "Food insecurity raw value", 
                     "Drug overdose deaths raw value", 
                     "Median household income raw value", 
                     "% below 18 years of age raw value",
                     "% 65 and older raw value", 
                     "% Hispanic raw value", 
                     "% Females raw value", "% Rural raw value",
                     "Adult obesity raw value",
                     "Income inequality raw value",
                     "Uninsured adults raw value")

eco_health_covars <- names(analytic_data2020) %in% eco_vars_of_int

covars <- analytic_data2020[eco_health_covars]

# Merge case data with covarites
day25_covars <- merge(day25, covars, by = c("Name", "State"))

day25_covars$log_pop_dense <- log(day25_covars$Population/(day25_covars$LandArea))

```

Import Google mobility data 
```{r mobility data, message=FALSE, results="hide", warning=FALSE}

# Data 1
grocery_pharmacy <- read_csv("Data/Mobility/google-mobility-us-groceryAndPharmacy.csv")

df.1 <- 
  grocery_pharmacy %>%
  gather(key = date, value = value, -State)
df.1$type <- "grocery_pharmacy"


# Data 2
parks <- read_csv("Data/Mobility/google-mobility-us-parks.csv")

df.2 <- 
  parks %>%
  gather(key = date, value = value, -State)
df.2$type <- "parks"


# Data 3
residential <- read_csv("Data/Mobility/google-mobility-us-residential.csv")

df.3 <- 
  residential %>%
  gather(key = date, value = value, -State)
df.3$type <- "residential"


# Data 4
retailAndRecreation <- read_csv("Data/Mobility/google-mobility-us-retailAndRecreation.csv")

df.4 <- 
  retailAndRecreation %>%
  gather(key = date, value = value, -State)
df.4$type <- "retailAndRecreation"


# Data 5
transitStations <- read_csv("Data/Mobility/google-mobility-us-transitStations.csv")

df.5 <- 
  transitStations %>%
  gather(key = date, value = value, -State)
df.5$type <- "transitStations"


# Data 6
workplaces <- read_csv("Data/Mobility/google-mobility-us-workplaces.csv")

df.6 <- 
  workplaces %>%
  gather(key = date, value = value, -State)
df.6$type <- "workplaces"


mobility.data <-
  left_join(df.1, df.2, by = c("State", "date", "value", "type")) %>%
  left_join(df.3) %>%
  left_join(df.4) %>%
  left_join(df.5) %>%
  left_join(df.6)

mobility_data_long <- rbind(df.1, df.2,df.3,df.4,df.5,df.6)
mobility_data_wide <- spread(mobility_data_long, date, value)


```


Run Simple Triple Interaction Model and get Marginal Effects for Transportation


```{r running simple linear model}

## set up bootstrap CI function
 bootstrapCI <- function(model, perc, boot_pred_data, type) {
  nr <- nrow(model$data)
  data <- model$data
  new_data <- data[sample(1:nr, size = nr, replace = TRUE), ]
  up <- update(model, data = new_data)
  
  #norm <- sum(predict(up, newdata = new_data, type = 'response'), na.rm = TRUE)
  
  boot_pred_data[,"log_Pub_Trans"] <- boot_pred_data$log_Pub_Trans - (boot_pred_data$log_Pub_Trans*perc)
  
  if (type == 'count') {
      perc_red <- sum(predict(up, newdata = boot_pred_data, type = 'response'), na.rm=TRUE)
  } else {
    perc_red <- mean(predict(up, newdata = boot_pred_data, type = 'response'), na.rm=TRUE)
  }
  
  return(perc_red)
 }

#set up percentiles for trans reduction
percents <- c(0.0, 0.10, 0.20, 0.30, 0.40, 0.50, 0.60, 0.70, 0.80, 0.90) 

#log transformation because the data is skewed forday 25 only
day25$log_pop_dense <- log(day25$popland+1)
day25$log_GDP <- log(day25$GDP+1)
day25$log_Pub_Trans <- log(day25$CommutingByPublicTransportation + 1)

#log transformation because the data is skewed forday 25 covars - 8 regions reduced compared to day25 data only

day25_covars$log_pop_dense <- log(day25_covars$popland+1)
day25_covars$log_GDP <- log(day25_covars$GDP+1)
day25_covars$log_Pub_Trans <- log(day25_covars$CommutingByPublicTransportation + 1)


#models
est_day25_pois_model <- glm(ConfirmedCasesDay25~ log_pop_dense+log_GDP+log_Pub_Trans + 
                        log(Population) + 
                        `% Females raw value` + 
                        `% 65 and older raw value` + 
                        `Adult obesity raw value`+ 
                        `Physical inactivity raw value` + 
                        `Unemployment raw value` +
                        `Income inequality raw value` +
                        `Poor or fair health raw value`+
                        `Uninsured adults raw value` +
                        `Adult smoking raw value`,
                       family = "poisson", 
                       data = day25_covars, 
                       offset(log(Population)))

est_day_snc_t1_pois_model <- glm(DayOfFirstCases~ log_pop_dense+log_GDP+log_Pub_Trans + 
                        log(Population) + 
                        `% Females raw value` + 
                        `% 65 and older raw value` + 
                        `Adult obesity raw value`+ 
                        `Physical inactivity raw value` + 
                        `Unemployment raw value` +
                        `Income inequality raw value` +
                        `Poor or fair health raw value`+
                        `Uninsured adults raw value` +
                        `Adult smoking raw value`,
                       family = "gaussian", 
                       data = day25_covars)

summary(est_day25_pois_model)
summary(est_day_snc_t1_pois_model)
           

model_day25 <- est_day25_pois_model
model_snc_t1 <- est_day_snc_t1_pois_model

results_day25 <- as.data.frame(matrix(nrow = length(percents), ncol = 5))
colnames(results_day25) <- c('Percent Pub Trans Reduction',
                             'Boot Pred', 
                             'Boot Low' , 
                             'Boot High', 
                             'Init. Model Pred')

results_day_first <- as.data.frame(matrix(nrow = length(percents), ncol = 5))
colnames(results_day_first) <- c('Percent Pub Trans Reduction',
                                 'Boot Pred', 
                                 'Boot Low' , 
                                 'Boot High', 
                                 'Init. Model Pred')


for (i in 1:length(percents)) {
  
  perc <- percents[i]
  data_temp <- day25_covars
  
  data_temp$log_Pub_Trans <- day25_covars$log_Pub_Trans - (day25_covars$log_Pub_Trans*perc)
  
  pred_perc_red <- predict(model_day25, newdata = data_temp, type = 'response')
  pred_day_red <- predict(model_snc_t1, newdata = data_temp, type = 'response')
  
  sum_perc_red <- sum(pred_perc_red, na.rm = TRUE)
  mean_first_day <- mean(pred_day_red, na.rm = TRUE)
  
  boot_day25 <- replicate(1000, bootstrapCI(model = model_day25, 
                                      perc =  perc, 
                                      boot_pred_data = day25_covars,
                                      type = 'count'))
  
  boot_day1 <- replicate(1000, bootstrapCI(model = model_snc_t1, 
                                      perc =  perc, 
                                      boot_pred_data = day25_covars,
                                      type = 'day'))
  
  #mean_boot_diff <- mean(boot, na.rm = TRUE)
  
  CI_boot_day25 <- quantile(boot_day25, probs = c(0.025,0.50, 0.975), na.rm = TRUE)
  CI_boot_day1 <- quantile(boot_day1, probs = c(0.025,0.50, 0.975), na.rm = TRUE)

  
  results_day25[i,1] <- perc
  results_day25[i,2] <- CI_boot_day25[[2]]
  results_day25[i,3] <- CI_boot_day25[[1]]
  results_day25[i,4] <- CI_boot_day25[[3]]
  results_day25[i,5] <- sum_perc_red
  
  results_day_first[i,1] <- perc
  results_day_first[i,2] <- CI_boot_day1[[2]]
  results_day_first[i,3] <- CI_boot_day1[[1]]
  results_day_first[i,4] <- CI_boot_day1[[3]]
  results_day_first[i,5] <- mean_first_day
  
}

```

Plot cases over time marginally attributed to each reduction level - comparing bootstrap 50% percentile to initial model predictions as a sanity check of the bootstrap CI estimates: 

```{r plot counts day 25}

ggplot(results_day25, aes(x=`Percent Pub Trans Reduction`, y=`Boot Pred`)) + 
    geom_errorbar(aes(ymin=`Boot Low`, ymax=`Boot High`), width=.1) +
    geom_line() +
    geom_point()

ggplot(results_day25, aes(x=`Percent Pub Trans Reduction`, y=`Init. Model Pred`)) + 
    geom_errorbar(aes(ymin=`Boot Low`, ymax=`Boot High`), width=.1) +
    geom_line() +
    geom_point()



```

Plot average day since first case change due to public transit reduction, same method: 

```{r plot days since t1}

ggplot(results_day_first, aes(x=`Percent Pub Trans Reduction`, y=`Boot Pred`)) + 
    geom_errorbar(aes(ymin=`Boot Low`, ymax=`Boot High`), width=.1) +
    geom_line() +
    geom_point()

ggplot(results_day_first, aes(x=`Percent Pub Trans Reduction`, y=`Init. Model Pred`)) + 
    geom_errorbar(aes(ymin=`Boot Low`, ymax=`Boot High`), width=.1) +
    geom_line() +
    geom_point()

```


Setting Up SuperLearner and TMLE

Can include more covariates and not assuming linearity of the model - sanity check to make sure parametric model is sound:

```{r setup spec, eval=FALSE}
##set up node list

node_list <- list(
  W = c('popland', 'GDP',eco_vars_of_int[-c(1,2)]),
  A = "pub_trans_quantile",
  Y = "ConfirmedCasesDay25"
) 

#process missing
processed <- process_missing(day25_covars, node_list)
COVID_data <- processed$data
node_list <- processed$node_list

ate_spec <- tmle_ATE(
  treatment_level = 10,
  control_level = 1
)

sl3_list_learners("continuous")
sl3_list_learners("binomial")
sl3_list_learners("categorical")

# choose y learners 
lrnr_mean <- make_learner(Lrnr_mean)
lrnr_xgboost <- make_learner(Lrnr_xgboost)
Lrnr_glm <- make_learner(Lrnr_glm)
Lrnr_hal9001 <- make_learner(Lrnr_hal9001)

# choose a learners 
Lrnr_grf <- make_learner(Lrnr_grf)
#Lrnr_multivariate <- make_learner(Lrnr_multivariate)


# define metalearners appropriate to data types
ls_metalearner <- make_learner(Lrnr_nnls)
#mn_metalearner <- make_learner(Lrnr_solnp, metalearner_linear_multinomial,
                               #loss_loglik_multinomial)

sl_Y <- Lrnr_sl$new(learners = list(lrnr_mean, lrnr_xgboost, Lrnr_glm, Lrnr_hal9001),
                    metalearner = ls_metalearner)

sl_A <- Lrnr_sl$new(learners = list(Lrnr_grf))
                    
learner_list <- list(A = sl_A, Y = sl_Y)

```


```{r fit_TMLE, eval = FALSE}

tmle_fit <- tmle3(ate_spec, COVID_data, node_list, learner_list)
tmle_fit
estimates <- tmle_fit$summary$psi_transformed
estimates

```


Initial cross validated predictions: not used now 


```{r matrix, eval=FALSE}

States <- unique(day25$State)
CV.risk <- matrix(NA, nrow=length(States), ncol=6)

#estimates<- matrix(NA, nrow=500, ncol=4)
ObsData <- day25

for(i in States){
  
idx <- match(i,States)
  
validation_data <- ObsData %>% filter(State == i)
training_data <- ObsData %>% filter(State != i)

est1.model <- glm(ConfirmedCasesDay25 ~ popland + GDP + CommutingByPublicTransportation, family = "gaussian", data = training_data)

est2.model <- glm(ConfirmedCasesDay25 ~ popland + GDP + CommutingByPublicTransportation, family = "poisson", data = training_data)

est3.model <- glm(ConfirmedCasesDay25 ~ popland *GDP * CommutingByPublicTransportation, family = "gaussian", data = training_data)

est4.model <- glm(ConfirmedCasesDay25 ~ popland *GDP * CommutingByPublicTransportation, family = "poisson", data = training_data)

est5.model <- glm(ConfirmedCasesDay25 ~ popland + GDP , family = "gaussian", data = training_data)

est6.model <- glm(ConfirmedCasesDay25 ~ popland + GDP , family = "poisson", data = training_data)


predict.est1 <- predict(est1.model, newdata = validation_data)
predict.est2 <- predict(est2.model, newdata = validation_data)
predict.est3 <- predict(est3.model, newdata = validation_data)
predict.est4 <- predict(est4.model, newdata = validation_data)
predict.est5 <- predict(est5.model, newdata = validation_data)
predict.est6 <- predict(est6.model, newdata = validation_data)
#predict.est7 <- predict(est7.model, newdata = validation_data)

l2.hat1 <- mean((validation_data$ConfirmedCasesDay25 - predict.est1)^2)
l2.hat2 <- mean((validation_data$ConfirmedCasesDay25 - predict.est2)^2)
l2.hat3 <- mean((validation_data$ConfirmedCasesDay25 - predict.est3)^2)
l2.hat4 <- mean((validation_data$ConfirmedCasesDay25 - predict.est4)^2)
l2.hat5 <- mean((validation_data$ConfirmedCasesDay25 - predict.est5)^2)
l2.hat6 <- mean((validation_data$ConfirmedCasesDay25 - predict.est6)^2)
#l2.hat7 <- mean((validation_data$ConfirmedCasesDay25 - predict.est7)^2)

CV.risk[idx,] <- c(l2.hat1, l2.hat2, l2.hat3, l2.hat4,l2.hat5,l2.hat6)
} 

colnames(CV.risk) <- c("l2.est1", "l2.est2", "l2.est3", "l2.est4", "l2.est5", "l2.est6")
CV.risk <- as.data.frame(CV.risk)

CV.risk
mses <- colMeans(CV.risk, na.rm = TRUE)

match(mses, min(mses))
```

```{r modeling, eval=FALSE}
Y <- day25$ConfirmedCasesDay25
X <-  subset(day25, select= -ConfirmedCasesDay25)
X <- as.data.frame(X)

Q_lib <- c("SL.mean", "SL.glmnet", "SL.ranger", "SL.rpartPrune", "SL.bayesglm")
g_lib <- c("SL.mean", "SL.glmnet")


vim <- varimpact(Y = Y, data = X, Q.library = Q_lib, g.library = g_lib, family="gaussian" )
vim$results_all

plot_var("Population", vim)


```

Junk Drawer


```{r junk_drawer, eval = FALSE}

boot_log <- replicate(1000, bootstrapCI(model = est1.log_model, newdata = newdata))

day25$uci_log <- apply(boot_log, MARGIN = 1, FUN = quantile, probs = 0.925, na.rm=TRUE)
day25$lci_log <- apply(boot_log, MARGIN = 1, FUN = quantile, probs = 0.025, na.rm=TRUE)
day25$fit_log <- apply(boot_log, MARGIN = 1, FUN = quantile, probs = 0.5, na.rm=TRUE)

day25$uci <- apply(boot, MARGIN = 1, FUN = quantile, probs = 0.925, na.rm=TRUE)
day25$lci <- apply(boot, MARGIN = 1, FUN = quantile, probs = 0.025, na.rm=TRUE)
day25$fit <- apply(boot, MARGIN = 1, FUN = quantile, probs = 0.5, na.rm=TRUE)



g3_log <- ggplot(day25, aes(x = day25$CommutingByPublicTransportation, y = ConfirmedCasesDay25)) +
  theme_bw() +
  geom_point() +
  geom_line(aes(y = fit)) +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3)

g3_log

g3 <- ggplot(day25, aes(x = day25$CommutingByPublicTransportation, y = ConfirmedCasesDay25)) +
  theme_bw() +
  geom_point() +
  geom_line(aes(y = fit_log)) +
  geom_ribbon(aes(ymin = lci_log, ymax = uci_log), alpha = 0.3)

g3


#predictions for 10 and 90 quantiles log model
predict.log_est_10 <- predict(est1.log_model, newdata = marginal_data_10, type='response')
predict.log_est_90 <- predict(est1.log_model, newdata = marginal_data_90, type='response')

#predictions for 10 and 90 quantiles 
predict.est_10 <- predict(est1.model, newdata = marginal_data_10, type='response')
predict.est_90 <- predict(est1.model, newdata = marginal_data_90, type='response')

log_marg_10_90_diff <- mean(predict.log_est_90 - predict.log_est_10, na.rm = TRUE)
marg_10_90_diff <- mean(predict.est_90 - predict.est_10, na.rm = TRUE)
```

The marginal differences for public transportation at 90 vs. 10 percentile
```{r marg diff with conf int, eval = FALSE}
marg_10_90_diff
log_marg_10_90_diff
```






